---
title: "Reproducible Research Assignment I - Fitbit Analysis"
output: html_document
---

#Context of Study
This assignment uses data from a personal activity monitoring device. The device collects data at 5 minute intervals through out the day and  spans across two months of data from an anonymous individual collected during  months of October and November, 2012. The number of steps taken in 5 minute intervals are recorded each day.

The variables included in this dataset are:

-	steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

-	date: The date on which the measurement was taken in YYYY-MM-DD format

-	interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is a comma-separated-value (CSV) file and there are a total of 17,568 observations.


# Reading the dataset from working directory:

```{r}
tabular<-read.csv("~/R/Reproducible Research/repdata_data_activity/activity.csv",header=TRUE)
```

Printing summary stats
```{r}
summary(tabular)
```


# 1. Histogram of total number of steps taken each day

```{r}
tabular1<-aggregate(steps~ date,tabular,sum)
hist(tabular1$steps, xlab="Steps", main="Histogram -Total Steps Per Day")
```

# 2. Mean and Median of Initial Dataset

```{r}
mean(tabular1$steps)
median(tabular1$steps)
```


# 3. Time Series Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
tabular6<-aggregate(steps~interval,tabular,mean)
plot(tabular6$interval,tabular6$steps, type='l', xlab="Interval", ylab="Steps")
```


# 4. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
tabular6[which.max(tabular6$steps),]

```
As observable, 835 interval contains the maximum number of steps.

# 5. Imputing missing values
There are a number of days/intervals where there are missing values (coded as  NA ). The presence of missing days may introduce bias into some calculations or summaries of the data. Calculating and reporting the total number of missing values in the dataset (i.e. the total number of rows with  NA s)

Counting NA's in Dataset>Column 'steps'
```{r}
sum(is.na(tabular$steps))

```

Filling the missing values with the mean for interval and creating new dataset

```{r}
library(plyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
tabular_new <- ddply(tabular, ~ interval, transform, steps = impute.mean(steps))
tabular_aggr<-aggregate(steps~ date,tabular_new,sum)

```

# 6. Histogram of the total number of steps taken each day (with filled in NA's)

```{r}
hist(tabular_aggr$steps, xlab="Steps", main="Histogram -Total Steps Per Day")
```

#7. Mean and Median for filled in dataset
```{r}
mean(tabular_aggr$steps)
median(tabular_aggr$steps)
```

- Do these values differ from the estimates from the first part of the assignment? 
Yes, median has changed.
```{r}

```


- What is the impact of imputing missing data on the estimates of the total daily number of steps?
There is no change in mean but median.


#8.Are there differences in activity patterns between weekdays and weekends?

Creating a new factor variable 'type' in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
tabular_new$date<-as.POSIXct(tabular_new$date)
tabular_new$day<-weekdays(tabular_new$date)
tabular_new$type<-ifelse(tabular_new$day %in% ("Sunday"),"Weekend", ifelse(tabular_new$day %in% ("Saturday"),"Weekend","Weekday"))
tabular_new2<-aggregate(steps~ interval+type,tabular_new,mean)
```

#9. Panel Time-Series Plot (Steps vs interval) Facet by Day Type
Panel plot containing a time series plot (i.e.  type = "l" ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
library(ggplot2)
ggplot(tabular_new2, aes(x = interval, y = steps)) +
 geom_line() + facet_grid(type ~ . )

```
